Characteristic polynomial of p-adic matrices

### 2017-01-18 charpoly/charpoly.tex
Begun bibliography

\begin{prop} 
Algorithm 1 computes $H$ and $P$ realizing an approximate Hessenberg form of $M.$
$P$ is exact over finite extensions of $\mathbb{Q}_p$ and $k((X))$, and the computation is in $O(n^3)$ operations in $K$ at precision the maximum precision of a coeffient in $M.$
\end{prop}
\begin{proof}
Let us assume that $K$ is a finite extensions of $\mathbb{Q}_p$ or $k((X)).$
Inside the nested \textbf{for} loop, if we want to eliminate $p^{u_y} \varepsilon_y+O(p^{n_y})$ with pivot $p^{u_x} \varepsilon_x+O(p^{n_x}),$
with the $\varepsilon$'s being units,
the corresponding coefficient of the corresponding shear matrix is the lift (in $\mathbb{Z}, $  $k[X]$ or adequate extension) of $p^{u_y-u_x} \varepsilon_y \varepsilon_x^{-1} \mod \pi^{\min (n_x,n_y)}.$
Exactness follows directly. Over other fields, we can not lift, but the computations are still valid.
The rest is clear.
\end{proof}

\begin{rem} \label{rem:char_pol_from_hessenberg}
From an Hessenberg form of $M,$ it is well known
that one can compute the characteristic polynomial of 
$M$ in $O(n^3)$ operations in $K.$ See \todo{cite Cohen} page
55 and 56.
However, this computation involves some divisions, and its
behaviour regarding to precision is not easy to
quantify.
\end{rem}

@@

\begin{prop}
Let $cond(M)$ be the valuation of the last invariant factor of $M.$
If $dM$ is a flat precision of $O(\pi^N)$ on $M$ then $M^{-1}$
can be computed at precision $O(\pi^{N-2cond(M)})$ by a \textbf{SNF} computation
and this lower-bound is optimal
at least when $N$ is large.
\end{prop}
\begin{proof}
The smallest valuation of a coefficient of $M^{-1}$ is $-cond(M).$
It is $-2cond(M)$ for $M^{-2}$ and it is then clear that $-2cond(M)$
can be obtained as the valuation of a coefficient of $du(M) \cdot dM$
and the smallest that can be achieved this way for $dM$ in a precision lattice
of flat precision. Hence the optimality of the bound given, at least when 
$N$ is large, thanks to \todo{Reference for precision lemma}.

Now, the computation of an SNF form over a CDVF was described in \todo{reference these Vaccon}.
From $M$ known at flat precision $O(\pi^N),$ we can obtain an exact $\Delta$ and $P,Q$ 
known at precision at least $O(\pi^{N-cond(M)})$ with coefficients in $O_K$
and determinant in $O_K^*$ realizing an SNF form of $M.$
There is no loss in precision when computing $P^{-1}$ and $Q^{-1}.$
Now, computing $\Delta^{-1}$ exactly, its smallest coefficient being  of valuation $-cond(M),$
we obtain $M^{-1}=Q^{-1} \Delta^{-1} P^{-1}$ known at precision at least $O(\pi^{N-2cond(M)}),$
which concludes the proof.
\end{proof}

@@

>>>>

\begin{prop} 
Algorithm 1 computes $H$ and $P$ realizing an approximate Hessenberg form of $M.$
$P$ is exact over finite extensions of $\mathbb{Q}_p$ and $k((X))$, and the computation is in $O(n^3)$ operations in $K$ at precision the maximum precision of a coefficent in $M.$
\end{prop}
\begin{proof}
Let us assume that $K$ is a finite extensions of $\mathbb{Q}_p$ or $k((X)).$
Inside the nested \textbf{for} loop, if we want to eliminate $p^{u_y} \varepsilon_y+O(p^{n_y})$ with pivot $p^{u_x} \varepsilon_x+O(p^{n_x}),$
with the $\varepsilon$'s being units,
the corresponding coefficient of the corresponding shear matrix is the lift (in $\mathbb{Z}, $  $k[X]$ or adequate extension) of $p^{u_y-u_x} \varepsilon_y \varepsilon_x^{-1} \mod \pi^{\min (n_x,n_y)}.$
Exactness follows directly. Over other fields, we can not lift, but the computations are still valid.
The rest is clear.
\end{proof}

\begin{rem} \label{rem:char_pol_from_hessenberg}
From an Hessenberg form of $M,$ it is well known
that one can compute the characteristic polynomial of 
$M$ in $O(n^3)$ operations in $K.$ See \cite{Cohen:2013} page
55 and 56.
However, this computation involves some divisions, and its
behaviour regarding to precision is not easy to
quantify.
\end{rem}

@@

\begin{prop}
Let $cond(M)$ be the valuation of the last invariant factor of $M.$
If $dM$ is a flat precision of $O(\pi^N)$ on $M$ then $M^{-1}$
can be computed at precision $O(\pi^{N-2cond(M)})$ by a \textbf{SNF} computation
and this lower-bound is optimal
at least when $N$ is large.
\end{prop}
\begin{proof}
The smallest valuation of a coefficient of $M^{-1}$ is $-cond(M).$
It is $-2cond(M)$ for $M^{-2}$ and it is then clear that $-2cond(M)$
can be obtained as the valuation of a coefficient of $du(M) \cdot dM$
and the smallest that can be achieved this way for $dM$ in a precision lattice
of flat precision. Hence the optimality of the bound given, at least when 
$N$ is large, thanks to Lemma 3.14 of \cite{CRV:2014}

Now, the computation of an SNF form over a CDVF was described in \cite{Vaccon-these}.
From $M$ known at flat precision $O(\pi^N),$ we can obtain an exact $\Delta$ and $P,Q$ 
known at precision at least $O(\pi^{N-cond(M)})$ with coefficients in $O_K$
and determinant in $O_K^*$ realizing an SNF form of $M.$
There is no loss in precision when computing $P^{-1}$ and $Q^{-1}.$
Now, computing $\Delta^{-1}$ exactly, its smallest coefficient being  of valuation $-cond(M),$
we obtain $M^{-1}=Q^{-1} \Delta^{-1} P^{-1}$ known at precision at least $O(\pi^{N-2cond(M)}),$
which concludes the proof.
\end{proof}

@@

\bibliographystyle{plain}
\bibliography{charpoly}
